

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Problems and Applications}
\exercise{Stock \& Watson, Introduction (4th), Chapter~4, Empirical Exercise~1.}
Get the data file \textbf{Growth}, which contains data on average growth rates from 1960 through 1995 for 65 countries, along with variables that are potentially related to growth. A detailed description is given in \textbf{Growth\_Description}. In this exercise, you will investigate the relationship between growth and trade.
\begin{enumerate}
\item Construct a scatterplot of average annual growth rate ($Growth$) on the average trade share ($TradeShare$). Does there appear to be a relationship between the variables?
\item One country, Malta, has a trade share much larger than the other countries. Find Malta on the scatterplot. Does Malta look like an outlier?
\item Using all observations, run a regression of Growth on $TradeShare$. What is the estimated slope? What is the estimated intercept? Use the regression to predict the growth rate for a country with a trade share of $0.5$ and for another with a trade share equal to $1.0$.
\item Estimate the same regression, excluding the data from Malta. Answer the same questions in (c).
\item Plot the estimated regression functions from (c) and (d). Using the scatterplot in (a), explain why the regression function that includes Malta is steeper than the regression function that excludes Malta.
\item Where is Malta? Why is the Malta trade share so large? Should Malta be included or excluded from the analysis?
\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

