

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Problems and Applications}
\exercise{Stock \& Watson, Introduction (4th), Chapter~4, Exercise~3.}
A regression of average weekly earnings (AWE, measured in dollars) on age (measured in years) using a random sample of college-educated full-time workers aged $25$–$65$ yields the following:
\begin{align*}
\verywidehat{AWE} 
  = 696.7 + 9.6 \times Age,
  \quad R^2 = 0.023,
  \quad \vn{SER} = 624.1
\end{align*}
\begin{enumerate}
\item Explain what the coefficient values $696.7$ and $9.6$ mean.
\item The standard error of the regression ($\vn{SER}$) is $624.1$. What are the units of measurement for the $\vn{SER}$? (Dollars? Years? Or is $\vn{SER}$ unit free?)
\item The regression $R^2$ is $0.023$. What are the units of measurement for the $R^2$? (Dollars? Years? Or is $R^2$ unit free?)
\item What does the regression predict will be the earnings for a 25-year-old worker? For a 45-year-old worker?
\item Will the regression give reliable predictions for a 99-year-old worker? Why or why not?
\item Given what you know about the distribution of earnings, do you think it is plausible that the distribution of errors in the regression is normal? (Hint: Do you think that the distribution is symmetric or skewed? What is the smallest value of earnings, and is it consistent with a normal distribution?)
\item The average age in this sample is $41.6$ years. What is the average value of AWE in the sample?
\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

