

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Summary}
\begin{itemize}
\item The population regression line, $\beta_0 + \beta_1 X$, is the mean of $Y$ as a function of the value of $X$. 
\item The slope, $\beta_1$, is the expected difference in $Y$ values between two observations with $X$ values that differ by one unit. 
\item The intercept, $\beta_0$, determines the level of the regression line.
\item The population regression line can be estimated using sample observations $(Y_i, X_i)$ by ordinary least squares (OLS) or another suitable method. 
\item The OLS estimators of the regression intercept and slope are denoted $\hat{\beta}_0$ and $\hat{\beta}_1$. 
\item The predicted value of $Y$ given $X$ is $\hat{\beta}_0+\hat{\beta}_1X$.
\item The $R^2$ and standard error of the regression ($\vn{SER}$) are measures of how
close the values of $Y_i$ are to the estimated regression line. The $R^2$ is between $0$ and $1$, with a larger value indicating that the $Y_i$s are closer to the line. 
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Summary}
\begin{itemize}
\item A regression error is the deviation of the observed value from the true value. A regression residual is the difference between the observed value and the estimated value. The standard error of the regression estimates the standard deviation of the regression errors.
\item There are three key assumptions for estimating causal effects using the linear regression model: 
\begin{enuminline}
\item The regression errors, $u_i$, have a mean of $0$, conditional on the regressors $X_i$: $\exp[u_{i}|X_{i}]=0$. 
\item The sample observations are i.i.d. random draws from the population; \item Large outliers are unlikely: $X_{i}$ and $Y_{i}$ have nonzero finite fourth moments.
\end{enuminline}
If these assumptions hold, the OLS estimator $\hat{\beta}_1$ is (1) an unbiased estimator of the causal effect $\beta_1$, (2) consistent, and (3) normally distributed when the sample is large.
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

