---
title: "Empirical Exercise 1, Stock and Watson Chapter 3"
subtitle: "Econ 440 - Introduction to Econometrics"
author: "Your Name + CWID + youremail@fullerton.edu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: default
---

```{r setup,  include=FALSE, eval=TRUE}
library(ggplot2)
library(dplyr)
library(tidyr)
options(digits=5)
```


### Tips
Explicitly mark the question or question number in your code. Show the output, not just the code: A few months/years from now, the packages will have been updated and the code may no longer run, so it's a good habit to keep a record of the output. Don't forget to answer the questions! 

### Load dataset
```{r data, include=TRUE, eval=TRUE}
library(readxl)
df <- read_xlsx("CPS96_15.xlsx", trim_ws=TRUE)
head(df)
```


## (a) 

### (i) 

Compute the sample mean for average hourly earnings (AHE) in 1996 and 2015.

```{r,  include=TRUE, eval=TRUE}
mu.1996 = mean(df[df$year == 1996,]$ahe)
## 12.693
mu.2015 = mean(df[df$year == 2015,]$ahe)
## 21.237
```

### (ii) 

Compute the sample standard deviation for AHE in 1996 and 2015.
```{r,  include=TRUE, eval=TRUE}
sd.1996 = sd(df[df$year == 1996,]$ahe)
## 6.359
sd.2015 = sd(df[df$year == 2015,]$ahe)
## 12.125
```

### (iii) 

Construct a 95% confidence interval for the population means of AHE in 1996 and 2015.
```{r,  include=TRUE, eval=TRUE}
alpha = 0.05
n = length(df$year == 1996)
t = qt(1-alpha/2, n-1)
se = sd.1996/sqrt(n)
me = t*se
ci = c(mu.1996-me, mu.1996+me)
ci
# 12.585 12.802
```

### (iv) 

Construct a 95% confidence interval for the change in the population means of AHE between 1996 and 2015.
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (b) 

In 2015, the value of the Consumer Price Index (CPI) was 237.0. In 1996, the value of the CPI was 156.9. Repeat (a), but use AHE measured in real 2015 dollars ($2015); that is, adjust the 1996 data for the price inflation that occurred between 1996 and 2015.
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (c) 

If you were interested in the change in workers' purchasing power from 1996 to 2015, would you use the results from (a) or (b)? Explain.
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (d) 

Using the data for 2015:

### (i) 

Construct a 95% confidence interval for the mean of AHE for high school graduates.
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

### (ii) 

Construct a 95% confidence interval for the mean of AHE for workers with a college degree. 
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

### (iii) 

Construct a 95% confidence interval for the difference between the two means.
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (e) 

Repeat (d) using the 1996 data expressed in $2015.
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (f) 

Using appropriate estimates, confidence intervals, and test statistics, answer the following questions:

### (i) 

Did real (inflation-adjusted) wages of high school graduates increase from 1996 to 2015?
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

### (ii) 

Did real wages of college graduates increase?
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

### (iii) 

Did the gap between earnings of college and high school graduates increase? Explain.
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

