---
title: "Regression with R: Extract, Tabulate, Plot"
subtitle: "Econ 440 - Introduction to Econometrics"
author: "Patrick Toche, ptoche@fullerton.edu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: default
  pdf_document: default
---

```{r setup,  include=FALSE, eval=TRUE}
library(ggplot2)
library(broom)
library(dplyr)
library(tidyr)
library(ggthemes)
options(digits=5)
```

# Extract, Tabulate, Plot
In this notebook we explore how to extract regression coefficients, standard errors, p-values and other pieces of information contained in a linear regression model. We also learn how to augment a regression model, how to tabulate regression results, and how to plot regression lines.

## Dataset on Earnings
The data file **CPS2015** contains data for full-time, full-year workers, ages 25–34, with a high school diploma or B.A./B.S. as their highest degree. A detailed description is given in **CPS2015_Description**. 

```{r data, include=TRUE, eval=TRUE}
library(readxl)
df <- read_xlsx("CPS2015.xlsx", trim_ws=TRUE)
head(df)
```

## A Log-quadratic model
This nonlinear regression model is analyzed in Chapter 8 of Stock and Watson's **Introduction to Econometrics**. Run a linear model with $lm()$ to produce a model object, then call the $summary()$ function to extract basic information. 
```{r,  include=TRUE, eval=TRUE}
m1 <- lm(log(ahe) ~ age + I(age^2) + bachelor + female, data=df)
summary(m1)
```

## Look inside a model object
See inside the model object: a list that holds useful results obtained from the regression, including the estimates, residuals, and standard errors.
```{r,  include=TRUE, eval=TRUE}
str(m1)
```

You may save the coefficients to a list:
```{r,  include=TRUE, eval=TRUE}
coefs <- coef(m1)
names(coefs)
```

You may extract the coefficients by name:
```{r,  include=TRUE, eval=TRUE}
coefs["(Intercept)"]
coefs["bachelor"]
```
Remember that you can always invoke $str()$ on an object to examine its content and thus figure out how to extract its elements. For instance, if you call $str(coefs)$ you will find that the regression coefficient on $age$ is called "poly(age, 2, raw = TRUE)1", while the coefficient on $age^2$ is called "poly(age, 2, raw = TRUE)2". 

Beware: In strings, spaces and cases matter, so the following won't work!
```{r,  include=TRUE, eval=TRUE}
coefs["(intercept)"]
```

The $broom$ package offers a more convenient and more versatile interface to extract and transform the regression data.

## Explore the model object with broom
The $broom$ package provides convenience functions, including $tidy()$, $glance()$ and $augment()$. These functions always return a `tibble` (a modern dataframe).

Extract coefficients: $tidy()$ returns a nicely formatted dataframe:
```{r,  include=TRUE, eval=TRUE}
tidy(m1)
```

Same with pipe operator and a slice to select a coefficient of interest. The function $pull()$ is then used to extract the desired value, e.g. $pull(std.error)$. The result can be manipulated further, e.g. rounding:
```{r,  include=TRUE, eval=TRUE}
m1 %>% tidy() %>% slice(2) %>% pull(std.error) %>% round(.,3)
```

You can add confidence intervals with the $conf.int$ argument and extract the desired values with $pull()$:
```{r,  include=TRUE, eval=TRUE}
m1 %>% tidy(conf.int=TRUE, conf.level=0.80) %>% pull(conf.low)
```

The $augment()$ function can be used to extract the fitted values and residuals for the original observations. In the augmented dataframe/tibble, each of the new columns begins with a dot, e.g. `.fitted`, to avoid accidentally overwriting existing variable names. This data could be extracted and used to plot a regression line and confidence interval, for instance.
```{r,  include=TRUE, eval=TRUE}
ma <- augment(m1, data=df)
names(ma)
```

The $glance()$ function can be used to extract summary statistics for the entire regression, including the R-squared and the F-statistic. 
```{r,  include=TRUE, eval=TRUE}
mg <- glance(m1, data=df)
names(mg)  # glancing the regression statistics
```
For instance, the R-squared can be extracted with:
```{r,  include=TRUE, eval=TRUE}
R2 <- mg$r.squared  # or mg[["r.squared"]]
print(R2)
```

## Visualize the regression coefficients
The following plot allows you to quickly visualize the regression coefficients:
```{r,  include=TRUE, eval=TRUE}
m1 %>% tidy(conf.int=TRUE, conf.level=0.80)  %>%
  ggplot(., aes(estimate, term, xmin=conf.low, xmax=conf.high, height=0)) +
  geom_point() +
  geom_vline(xintercept=0, lty=4) +
  geom_errorbarh()
```

# Plot regression lines
```{r,  include=TRUE, eval=TRUE}
augment(m1, df, interval="confidence") %>%
  ggplot(data=.) +
  aes(x=age, 
      y=.fitted,
      group=interaction(-female, bachelor), 
      color=interaction(-female, bachelor)) +
    geom_line(size=1.5) -> p0
p0
```

## Fix the legend, select colors, tweak labels
```{r,  include=TRUE, eval=TRUE}
labs <- c("Female & High School", "Male & High School", "Female & Bachelor", "Male & Bachelor")  # create labels
p0 + scale_x_continuous(breaks=seq(25,35,1)) +
  scale_color_manual(name="",labels=labs, values=c("red", "forestgreen", "orange", "blue")) +
  guides(color=guide_legend(reverse=TRUE)) +
  ylab("fitted ahe") +
  ggtitle("log-quadratic model") -> p1
p1
```
## Add confidence intervals
```{r,  include=TRUE, eval=TRUE}
p1 + geom_ribbon(aes(ymin=.lower, ymax=.upper), alpha=.25, color=NA) -> p2
p2 + theme_minimal()
```

## Add theme to the plot
```{r,  include=TRUE, eval=TRUE}
library(ggthemes)  # to theme the plot
p2 + theme_wsj(base_size=10)
```

## Confidence Intervals for Parameters
The `dplyr` package (part of `tidyverse`) contains a convenient $full_join()$ function that may be used to merge dataframes containing estimates calculated with `broom` functions like $tidy()$ and $glance()$. First, create a dataframe which contains both the data estimates (using $tidy()$) and the model's summary statistics (using $glance()$). Then use $pull()$ to get the desired statistics.
```{r,  include=TRUE, eval=TRUE}
dplyr::full_join(
  df %>% group_modify(.f = ~ tidy(m1, conf.int=TRUE, conf.level=0.99)),
  df %>% group_modify(.f = ~ glance(m1))
) -> dm
```

Then pull the desired statistics:
```{r,  include=TRUE, eval=TRUE}
dm %>% pull(estimate)
dm %>% pull(conf.low)
dm %>% pull(conf.high)
```

## Multiple models
Let's compute several regression models.
```{r,  include=TRUE, eval=TRUE}
m2 <- lm(log(ahe) ~ age + I(age^2) + I(age^3) + bachelor + female, data=df)
m3 <- lm(log(ahe) ~ log(age) + bachelor + female, data=df)
```

## Fitted values and residuals
Clusters of points can potentially cause the errors to be heteroskedastic. To visualize this, we color the points according to the four combinations of Male/Female and Bachelor/High-School. We suppress the legend for clarity and set the color palette with the $scale_color_brewer()$ function of the ``ggplot2`` package. 
```{r,  include=TRUE, eval=TRUE}
augment(m1, df, interval="confidence") %>%
  ggplot(data=., aes(x=.fitted, y=.resid)) + 
  geom_point(aes(color=interaction(-female, bachelor))) +
  geom_smooth(method="lm") +
  scale_color_brewer(palette="Set1") +
  theme_minimal() +
  theme(legend.position="none")
```

Look at the distribution of the residuals for evidence of heteroskedasticity:
```{r,  include=TRUE, eval=TRUE}
augment(m1, df, interval="confidence") %>%
  ggplot(data=., aes(x=.resid)) + 
  geom_histogram(color="white", fill="cornflowerblue") +
  theme_minimal()
```
No evidence of heteroskedasticity! However, the clustering suggests that we must use robust standard errors.

## Compute robust standard errors
The ``sandwich`` package computes robust Heteroscedasticity-Consistent Covariance estimators with the function $vcovHC()$. The $type$ argument can be used to specify estimators from $HC0$ (White's estimator) to $HC5$ (various refinements). The ``lmtest`` package provides a convenient function, $coeftest()$, to calculate the t test based on the variance-covariance matrix provided in the $vcov$ argument. A robust test may be computed as follows:
```{r,  include=TRUE, eval=TRUE}
library(sandwich)
library(lmtest)  # coeftest, waldtest
coeftest(m1, vcov=vcovHC(m1, type="HC1"))
```

## Compute robust Wald test
A heteroskedasticity robust F test may be calculated with the $waldtest$ function from package $lmtest$ (using White standard errors):
```{r,  include=TRUE, eval=TRUE}
m1.unrestricted <- lm(log(ahe) ~ age + I(age^2) + I(age^3) + bachelor + female, data=df)
waldtest(m1, m1.unrestricted, vcov=vcovHC(m1.unrestricted, type="HC0"))
```


# Tabulate Regression results
Tabulating regression results may be automated with the help of several packages. A popular solution is ``stargazer``. Here instead I use a newer package, ``texreg``.

## Output the table to screen with $screenreg()$:
```{r,  include=TRUE, eval=TRUE}
suppressMessages(library(texreg))
screenreg(list(m1, m2, m3))
```
## Export the table to the LaTeX format with $texreg()$:
```{r,  include=TRUE, eval=TRUE}
# htmlreg(list(m1, m2, m3), doctype = FALSE, star.symbol = "\\*")
texreg(list(`(1)`=m1, `(2)`=m2, `(3)`=m3), booktabs=TRUE, dcolumn=TRUE)
```

## Export table to HTML format with $htmlreg()$
[Not shown as it messes up the PDF output]

## Export table to PDF format
(the texreg argument ``use.packages=FALSE`` is set to suppress any package loading instructions in the preamble)
```{r, results='asis', echo=TRUE}
texreg(list(`(1)`=m1, `(2)`=m2, `(3)`=m3), table=FALSE, use.packages=FALSE)
```

Let's reorder the coefficients
```{r, results='asis', echo=TRUE}
texreg(list(`(1)`=m1, `(2)`=m2, `(3)`=m3), table=FALSE, use.packages=FALSE,
       reorder.coef = c(1,2,3,6,7,4,5))
```

