---
title: "The Dummy Variable Trap with R"
subtitle: "Econ 440 - Introduction to Econometrics"
author: "Patrick Toche, ptoche@fullerton.edu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup,  include=FALSE, eval=TRUE}
library(ggplot2)
library(broom)
library(dplyr)
library(tidyr)
options(digits=5)
```
```{r data, include=TRUE, eval=TRUE, echo=FALSE}
library(readxl)
df <- read_xlsx("caschool.xlsx", trim_ws=TRUE)
# Rename variables to match textbook notation
names(df)[names(df) == "testscr"] <- "TestScore"
names(df)[names(df) == "str"] <- "STR"
```
### Create regions:
To illustrate regression with binary variables, let's group California counties by geographical area, based on Census data. To keep the example simple, let's **arbitrarily** create 4 regions named: North, South, East, West, based on a cursory look at a map.
```{r,  include=TRUE, eval=TRUE, echo=FALSE}
# https://census.ca.gov/regions/

# Superior California: 
Superior.California <- c("Butte", "Colusa", "El Dorado", "Glenn", "Lassen", "Modoc", "Nevada", "Placer", "Plumas", "Sacramento", "Shasta", "Sierra", "Siskiyou", "Sutter", "Tehama", "Yolo", "Yuba")

# North Coast:
North.Coast <- c("Del Norte", "Humboldt", "Lake", "Mendocino", "Napa", "Sonoma", "Trinity")

# San Francisco Bay area
San.Francisco.Bay.Area <- c("Alameda", "Contra Costa", "Marin", "San Francisco", "San Mateo", "Santa Clara", "Solano")

# Northern San Joaquin Valley
Northern.San.Joaquin.Valley <- c("Alpine", "Amador", "Calaveras", "Madera", "Mariposa", "Merced", "Mono", "San Joaquin", "Stanislaus", "Tuolumne")

# Central Coast
Central.Coast <- c("Monterey", "San Benito", "San Luis Obispo", "Santa Barbara", "Santa Cruz", "Ventura")

# Southern San Joaquin Valley
Southern.San.Joaquin.Valley <- c("Fresno", "Inyo", "Kern", "Kings", "Tulare")

# Inland Empire
Inland.Empire <- c("Riverside", "San Bernardino")
 
# Los Angeles County
Los.Angeles.County <- c("Los Angeles")

# Orange County
Orange.County <- c("Orange")

# San Diego - Imperial
San.Diego.Imperial <- c("Imperial", "San Diego")

North.Counties <- c(Superior.California, North.Coast, San.Francisco.Bay.Area, Northern.San.Joaquin.Valley)
West.Counties <- Central.Coast
East.Counties <- Southern.San.Joaquin.Valley
South.Counties <- c(Inland.Empire, Los.Angeles.County, Orange.County, San.Diego.Imperial)
```

For instance, here is what the list $North.Counties$ looks like. 
```{r,  include=TRUE, eval=TRUE}
North.Counties
```
To view the complete code, see the ``Rmd`` file.

### Create dummy variables for each region:
```{r,  include=TRUE, eval=TRUE}
df$North <- df$county %in% North.Counties
df$East <- df$county %in% East.Counties
df$South <- df$county %in% South.Counties
df$West <- df$county %in% West.Counties
head(df[c("North", "East", "South", "West")], 10)
```

### Check consistency of the categories:
```{r,  include=TRUE, eval=TRUE}
unique(df$North + df$East + df$South + df$West) == 1
```

### Create a categorical variable for all 4 regions:
```{r,  include=TRUE, eval=TRUE}
df$Region <- NA
dummies <- c("North", "East", "South", "West")
for (col in dummies)
  df$Region[which(df[,col] == TRUE)] <- col
head(df$Region, 10)
```

# Regression With Categorical Variables
If you estimate a linear regression with all categories and an intercept, R will automatically drop one of the categories:
```{r,  include=TRUE, eval=TRUE}
m1 <- lm(TestScore ~ STR + Region, data=df)
summary(m1)
```
The categorical variable $Region$ stores the categories as strings (aka characters). When running the regression, R temporarily creates factor variables using alphabetical ordering, thus creating categories for $RegionNorth$, $RegionSouth$ and $RegionWest$, while omitting the first category in the list, $RegionEast$. 

To use, say, $North$, as the reference region, create factors in the desired order, and run the regression again:
```{r,  include=TRUE, eval=TRUE}
df$Region2 <- factor(df$Region, levels=c("North", "East", "South", "West"))
m2 <- lm(TestScore ~ STR + Region2, data=df)
summary(m2)
```

If the categories are already stored as factors but do not appear in the desired order, you can reset the factor levels:
```{r,  include=TRUE, eval=TRUE}
df$Region3 <- relevel(df$Region2, ref="South")
m3 <- lm(TestScore ~ STR + Region3, data=df)
summary(m3)
```



