% !TEX root = ../../MakeBeamer.tex
\title[Earnings]{Linear Regression: Earnings}
\date{}


\begin{document}


\inputfile{../Section-Cover-Problems}


\def\ask{\begin{align*}
\verywidehat{AWE} 
  = 696.7 + 9.6 \times Age,
  \quad R^2 = 0.023,
  \quad \vn{SER} = 624.1
\end{align*}\vspace*{-3ex}}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Problems and Applications}
\exercise{Stock \& Watson, Introduction (4th), Chapter~4, Exercise~3.}
A regression of average weekly earnings (AWE, measured in dollars) on age (measured in years) using a random sample of college-educated full-time workers aged $25$–$65$ yields the following:
\ask
\begin{enumerate}
\item Explain what the coefficient values $696.7$ and $9.6$ mean.
\item The standard error of the regression ($\vn{SER}$) is $624.1$. What are the units of measurement for the $\vn{SER}$? (Dollars? Years? Or is $\vn{SER}$ unit free?)
\item The regression $R^2$ is $0.023$. What are the units of measurement for the $R^2$? (Dollars? Years? Or is $R^2$ unit free?)
\item What does the regression predict will be the earnings for a 25-year-old worker? For a 45-year-old worker?
\item Will the regression give reliable predictions for a 99-year-old worker? Why or why not?
\item Given what you know about the distribution of earnings, do you think it is plausible that the distribution of errors in the regression is normal? (Hint: Do you think that the distribution is symmetric or skewed? What is the smallest value of earnings, and is it consistent with a normal distribution?)
\item The average age in this sample is $41.6$ years. What is the average value of AWE in the sample?
\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Problems and Applications}
\ask

\begin{enumerate}\setcounter{enumi}{0}

\item Explain what the coefficient values $696.7$ and $9.6$ mean.

\begin{answer}
The coefficient $696.7$ is the intercept of the regression line. The greater the intercept, the higher the overall position of the regression line. The intercept gives the mean value of $AWE$ when $Age=0$. 
The coefficient $9.6$ measures the marginal effect of $Age$ on $AWE$. On average, every increase of $Age$ by one year causes $AWE$ to increase by $\$9.6$. 
\end{answer}

\item The standard error of the regression ($\vn{SER}$) is $624.1$. What are the units of measurement for the $\vn{SER}$?

\begin{answer}
The $\vn{SER}$ is measured in the same units as the $y$ variable, that is dollars per week.
\end{answer}

\item The regression $R^2$ is $0.023$. What are the units of measurement for the $R^2$?

\begin{answer}
The $R^2$ is defined as a ratio, where the numerator and denominator are both measured in squared-dollars per week. The units cancel out and therefore $R^2$ is a pure number. 
\end{answer}

\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Problems and Applications}
\ask

\begin{enumerate}\setcounter{enumi}{3}

\item What does the regression predict will be the earnings for a $25$-year-old worker? For a $45$-year-old worker?

\begin{answer}
The regression predicts:
\begin{align*}
696.7 + 9.6 \times 25 & = 936.7 ~\text{for a $25$ year-old worker}\\ 
696.7 + 9.6 \times 45 & = 1,128.7 ~\text{for a $45$ year-old worker}
% 696.7+9.6*45
% 1128.7
% 696.7+9.6*25
% 936.7
\end{align*}
\end{answer}

\item Will the regression give reliable predictions for a $99$-year-old worker? Why or why not?

\begin{answer}
A regression may not give reliable predictions for out-of-sample $X$ values. In this case, $99$ years is quite far out of the sample, since the oldest person in the sample is $65$.
\end{answer}

\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Problems and Applications}
\ask

\begin{enumerate}\setcounter{enumi}{5}

\item Given what you know about the distribution of earnings, do you think it is plausible that the distribution of errors in the regression is normal? 

\begin{answer}
The distribution of earnings itself is not normally distributed; it is strongly positively skewed. 
\end{answer}

\item The average age in this sample is $41.6$ years. What is the average value of $AWE$ in the sample?

\begin{answer}
The mean $Age$ in the sample is $41.6$, from which we can compute the mean value of $AWE$ in the sample: 
\begin{align*}
\mean{AWE} 
    = 696.7 + 9.6 \times \mean{Age}
    = 696.7 + 9.6 \times 41.6 
    = 1,096.06
\end{align*}
\end{answer}

\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
