% !TEX root = ../../MakeBeamer.tex
\title[Test Scores]{Linear Regression: California Test Scores}
\date{}


\begin{document}


\inputfile{../Section-Cover-Problems}


\def\ask{\begin{align*}
\verywidehat{TestScore} 
  = 520.4 - 5.82 \times CS,
  \quad R^2 = 0.08,
  \quad \vn{SER} = 11.5
\end{align*}}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Test Scores}
\exercise{Stock \& Watson, Introduction (4th), Chapter~4, Exercise~1.}
A researcher, using data on class size (CS) and average test scores from $100$ third-grade classes, estimates the OLS regression:
\ask
\begin{enumerate}
\item A classroom has $22$ students. What is the regression's prediction for that classroom's average test score?
\item Last year a classroom had $19$ students, and this year it has $23$ students. What is the regression's prediction for the change in the classroom average test score?
\item The sample average class size across the $100$ classrooms is $21.4$. What is the sample average of the test scores across the $100$ classrooms?
\item What is the sample standard deviation of test scores across the $100$ classrooms?
\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Test Scores}
\ask

\begin{enumerate}\setcounter{enumi}{0}

\item A classroom has $22$ students. What is the regression's prediction for that classroom's average test score?

\begin{answer}
The regression's prediction is: 
\begin{align*}
\verywidehat{TestScore}_{|CS=22}
  & = 520.4 - 5.82 \times 22\\
  & = 392.36
% 520.4 - 5.82 * 22
% 392.36
\end{align*}
\end{answer}

\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Test Scores}
\ask

\begin{enumerate}\setcounter{enumi}{1}

\item Last year a classroom had $19$ students, and this year it has $23$ students. What is the regression's prediction for the change in the classroom average test score?

\begin{answer}
The regression's prediction is:
\begin{align*}
\Delta\verywidehat{TestScore}_{|CS=19\to23}
  & = (520.4 - 5.82 \times 23) - (520.4 - 5.82 \times 19)\\
  & = - 5.82 \times (23-19)\\
  & = -23.28
% -5.82*(23-19)
% -23.28
\end{align*}
The classroom size was increased by $4$ extra students, so we expect the average test score to fall by about $23$ points. 
\end{answer}

\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Test Scores}
\ask

\begin{enumerate}\setcounter{enumi}{2}

\item The sample average class size across the $100$ classrooms is $21.4$. What is the sample average of the test scores across the $100$ classrooms?

\begin{answer}
The sample average across the $100$ classrooms is:
\begin{align*}
\overline{TestScore}
  & = \hat{\beta}_{0} +  \hat{\beta}_{1} \times \overline{CS}\\
  & = 520.4 - 5.82 \times 21.4\\
  & = 395.852
% 520.4 - 5.82 * 21.4
% 395.852
\end{align*}
\end{answer}

\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Test Scores}
\ask

\begin{enumerate}\setcounter{enumi}{3}

\item What is the sample standard deviation of test scores across the $100$ classrooms?

\begin{answer}
The sample variance may be computed as the ratio of the total sum of squares to the residual sum of squares, $TSS/RSS$. 
The residual sum of squares, $RSS$, may be computed from the standard error of the regression, $\vn{SER}$:
\begin{align*}
\vn{SER} = \sqrt{\frac{\sum_{i=1}^{n}\hat{u}_{i}^2}{n-2}}
    = \sqrt{\frac{RSS}{n-2}}
\implies
RSS = (n-2) \vn{SER}^2
    = 98 \cdot 11.5^2
    = 12,961
\end{align*}
The total sum of squares, $TSS$, may be computed from the formula for $R^{2}$:
\begin{align*}
R^2 = \frac{ESS}{TSS}
    = 1 - \frac{RSS}{TSS}
\implies
TSS = \frac{RSS}{1-R^2}
    = \frac{12,961}{1-0.08^2}
    = 13,044
\end{align*}
And finally the variance of test scores:
\begin{align*}
s_{YY} 
    = \tfrac{1}{n-1}\sum_{i=1}^{n}(Y_{i}-\mean{Y})^2
    = \frac{TSS}{n-1}
    = \frac{13,044}{99}
    = 131.8
\implies
s_{Y} \approx 11.5 
% 13044/99 
% 131.7576
% sqrt(13044/99)
% 11.47857
\end{align*}
\end{answer}

\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
