---
title: 'Stock and Watson Chapter 8: Replication'
author: "Patrick Toche, ptoche@fullerton.edu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
  slidy_presentation:
    highlight: github
  beamer_presentation:
    latex_engine: xelatex
    slide_level: 2
    includes:
      in_header: setup.tex
    keep_tex: yes
  ioslides_presentation:
    highlight: github
subtitle: Econ 440 - Introduction to Econometrics
fontsize: 10pt
classoption: t, svgnames, handout
---


```{r setup,  include=FALSE, eval=TRUE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  paste0("\n \\", "scriptsize","\n\n", x, "\n\n \\normalsize")
})
knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      warning=FALSE)
library(ggplot2)
library(broom)
library(dplyr)
library(tidyr)
options(digits=5)
```

## Replication
Replicate regression results from James H. Stock and Mark W. Watson, Introduction to econometrics, Pearson, 4th Edition, Chapter 8. The data used is available in the Stata format $caschool.dta$ and in the Excel format $caschool.xlsx$.

## Load dataset 
Let's load the Stata dataset using the `haven` library.
```{r data, include=TRUE, eval=TRUE, echo=TRUE}
library(haven)
df <- read_dta("caschool.dta")
head(df)
```
You can also use the Excel file of course
```{r, include=TRUE, eval=FALSE, echo=TRUE}
library(readxl)
df <- read_xlsx("caschool.xlsx", trim_ws=TRUE)
```

## Data cleaning
The variable names are inconsistent with the textbook and other versions of the dataset used in other exercises, so let's rename them:
```{r, include=TRUE, eval=TRUE, echo=TRUE}
names(df)[names(df) == "avginc"] <- "income"
names(df)[names(df) == "testscr"] <- "testscore"
```

## Linear Regression of Test Score on District Income
### The simple linear regression.
```{r, include=TRUE, eval=TRUE, echo=TRUE}
m1 <- lm(testscore ~ income, data=df)
summary(m1)
```

## Tidy up with broom
Once we have estimated a model, it is convenient to use the `broom` library to tidy things up:
```{r, include=TRUE, eval=TRUE, echo=TRUE}
library(broom)
tidy(m1)
```

## Nonlinear Regression of Test Score on District Income
### The quadratic regression.
To run a nonlinear regression on $var$, we can use the $I()$ wrapper:
```{r, include=TRUE, eval=FALSE, echo=TRUE}
lm(testscore ~ income + I(income^2), data=df)
```
or use the more versatile $poly(var)$:
```{r, include=TRUE, eval=TRUE, echo=TRUE}
m2 <- lm(testscore ~ poly(income,2,raw=TRUE), data=df)
tidy(m2)
```
but note that we set the `raw=TRUE` option: ``if true, use raw and not orthogonal polynomials.''

## Nonlinear Regression of Test Score on District Income
### The cubic regression.
```{r, include=TRUE, eval=TRUE, echo=TRUE}
m3 <- lm(testscore ~ poly(income,3,raw=TRUE), data=df)
tidy(m3)
```

## Linear-Log Regression
### The linear-log regression.
```{r, include=TRUE, eval=TRUE, echo=TRUE}
m4 <- lm(testscore ~ log(income), data=df)
tidy(m4)
```

## Log-Linear Regression
### The linear-log regression.
```{r, include=TRUE, eval=TRUE, echo=TRUE}
m5 <- lm(log(testscore) ~ income, data=df)
tidy(m5)
```

## Log-Log Regression
### The log-log regression.
```{r, include=TRUE, eval=TRUE, echo=TRUE}
m6 <- lm(log(testscore) ~ log(income), data=df)
tidy(m6)
```
