% !TEX root = ../../MakeBeamer.tex
\title[Probability]{Review of Probability: Card Games}
\date{}


\begin{document}


\inputfile{../Section-Cover-Problems}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[allowframebreaks]%<1-| handout:0>[label=fancyframe]
\frametitle{Probability of a Hand}
\begin{emphbox}{}
Five cards are randomly selected from a standard deck. What is the probability of drawing the Four kings and the queen of hearts (in any order)? 
\end{emphbox}
\begin{itemize}
\item<2-> A standard problem of drawing without replacement! Which of these solutions is correct?
\item<3|only@3>\emph{Solution~1}
\dash The probability of drawing the $4$ kings is:
\begin{align*}
\frac{4}{52} \times \frac{3}{51} \times \frac{2}{50} \times \frac{1}{49} 
\end{align*}
\dash There is only one queen of hearts and $48$ cards left after the kings have been drawn. The probability of drawing $4$ kings and the queen of hearts is therefore:
\begin{align*}
\frac{4}{52} \times \frac{3}{51} \times \frac{2}{50} \times \frac{1}{49} \times \frac{1}{48} 
= \frac{24}{311875200}
\approx 0.000000077
\end{align*}

\item<4|only@4>\emph{Solution~2}
\dash There are $\dbinom{52}{5}$ ways of selecting $5$ cards from the deck.  
\dash There are exactly $\dbinom{4}{4}$ ways of selecting the $4$ kings. 
\dash Of the $48$ remaining cards, there are $\dbinom{48}{1}$ ways to select the fifth card. 
\dash The probability is therefore:
\begin{align*}
\dfrac{\binom{4}{4}\binom{48}{1}}{\binom{52}{5}} 
= \frac{1}{54145} 
\approx  0.000018
\end{align*}

\item<5|only@5>\emph{Solution~3}
\dash The probability of drawing $4$ kings and the queen of hearts is equal to the probability of selecting $5$ cards:
\begin{align*}
\frac{5}{52} \times \frac{4}{51} \times \frac{3}{50} \times \frac{2}{49} \times \frac{1}{48} 
= \frac{1}{2598960}
\approx 0.00000038
\end{align*}
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%\againframe<0| handout:1>{fancyframe}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Probability of a Hand}
\begin{itemize}
\item Recall that the number of order-free combinations is:
\begin{align*}
\binom{n}{k} = \comb{n}{k} = \frac{n!}{k!(n-k)!}
\end{align*}
\item Solution 1 gives the probability of drawing four kings first, in any order, and the queen of hearts in one given position. As there are $5$ positions the queen of hearts could be in, Solution 1 is too small by a factor of $5$. 
\item Solution 2 gives the probability of drawing four kings and any other card (not specifically the queen of hearts). Solution 2 is too large by a factor of $48$.
\item Solution 3 gives the correct probability.
There are $\binom{52}{5}$ ways to choose $5$ cards from the deck, and only one of those ways is the event of interest. The probability is therefore:
\begin{align*}
\frac{1}{\dbinom{52}{5}} 
= \frac{5! \times 47!}{52!}
\approx 0.00000038
\end{align*}
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Probability of a Hand}
\begin{itemize}
\item Equivalently, the correct answer can be broken down as follows: 
\begin{align*}
\frac{\binom{4}{4} \binom{1}{1}}{\binom{52}{5}}  = \frac{1}{\binom{52}{5}} 
\end{align*}
The numerator is the number of ways to choose $4$ kings from the $4$ available kings, $\binom{4}{4}$, multiplied by the number of ways to choose $1$ queen of heart from the only available queen of heart $\binom{1}{1}$. Thus, there is only $1$ such $5$-card hand.
\item The probability of drawing the four kings and the queen of hearts is therefore equal to the probability of drawing any \textit{specific} hand, e.g. ace of heart, three of diamonds, seven of clubs, nine of diamonds, jack of spades. 
\item The probability of drawing the four kings and any queen (whether hearts, diamonds, spades or clubs) is $4$ times as large (replace $\binom{1}{1}$ by $\binom{4}{1}$ in the formula above). 
\item The answers are related as follows:
\begin{align*}
\tfrac{1}{2598960}
& = ~ 5 ~ \times \tfrac{24}{311875200} \\
& = \tfrac{1}{48}\times \tfrac{1}{54145} 
\end{align*}
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% install.packages("arrangements")
% library("arrangements")
% ncombinations(52, 5)
% ## 2598960
% npermutations(52, 5)
% ## 311875200 


\end{document}
