---
title: "Empirical Exercise 1, Stock and Watson Chapter 2"
author: "Your Name + CWID + youremail@fullerton.edu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: default
subtitle: Econ 440 - Introduction to Econometrics
---

```{r setup,  include=FALSE, eval=TRUE}
library(ggplot2)
library(dplyr)
library(tidyr)
options(digits=5)
```


### Tips
Explicitly mark the question or question number in your code. Show the output, not just the code: A few months/years from now, the packages will have been updated and the code may no longer run, so it's a good habit to keep a record of the output. Don't forget to answer the questions! 

### Load dataset
```{r data, include=TRUE, eval=TRUE}
library(readxl)
df1 <- read_xlsx("Age_HourlyEarnings.xlsx",
                  col_names=TRUE,
                  skip=1,
                  trim_ws=TRUE)
```

### Remove empty rows and columns: simple and symmetric!
```{r,  include=TRUE, eval=TRUE}
df1 <- df1[, colMeans(is.na(df1)) != 1]
df1 <- df1[rowMeans(is.na(df1)) != 1,]
```
There are other ways to do this, e.g. using the `Filter()` and `complete.cases()` functions.

### Reshape data from wide to long form:
```{r,  include=TRUE, eval=TRUE}
library("tidyr")
df2 <- gather(df1, Age, Probability, -AHE)
```

### Make Age a numeric or integer value:
```{r,  include=TRUE, eval=TRUE}
df2$Age <- as.integer(df2$Age)
```
This is needed after conversion from wide (horizontal) to long (vertical), since the variable `Age` in dataframe `df2` is constructed from the column names in dataframe `df1`. And `colnames` are `strings` (aka `character vectors` in `R`).

### Now let's look at our data
```{r,  include=TRUE, eval=TRUE}
head(df2)
```

## (a) Compute the marginal distribution of Age.
This is a situation where data in wide format is convenient! Often data is more convenient in long format.

### Sum by column: The marginal distribution of Age
```{r,  include=TRUE, eval=TRUE}
colSums(df1[,names(df1) != "AHE"])
```

### Sum by row:
```{r,  include=TRUE, eval=TRUE}
rowSums(df1[,names(df1) != "AHE"])
```


If the data is in long format, we have to work a little harder. Below are several approaches.

### With the data in long format:
```{r,  include=TRUE, eval=TRUE}
sum(df2[df2$Age == 25,]$Probability)
sum(df2[df2$Age == 26,]$Probability)
sum(df2[df2$Age == 27,]$Probability)
```

### with a split/apply routine to group by Age
```{r,  include=TRUE, eval=TRUE}
sapply(split(df2, df2$Age), function(x) sum(x$Probability))
```

To see how this works, try this command: `split(df2, df2$Age)`. It returns a list of dataframes. Then `sapply` applies the function we defined as `function(x) sum(x$Probability)` to each dataframe.

### or we can always write a loop
```{r,  include=TRUE, eval=TRUE}
for (age in 25:34){
    print(sum(df2[df2$Age == age,]$Probability))
}
```

## (b) Compute the mean of AHE for each value of Age
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (c) Compute and plot the mean of AHE versus Age. Are average hourly earnings and age related? Explain.
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (d) Use the law of iterated expectations to compute the mean of AHE
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (e) Compute the variance of AHE.
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (f) Compute the covariance between AHE and Age.
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (g) Compute the correlation between AHE and Age.
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (h) Relate your answers in (f) and (g) to the plot you constructed in (c).
```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

