

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Course Overview}
\begin{itemize}
\item Economic theory suggests causal relationships and broad policy implications, but without measurement and quantification, economic theory lacks precision and real-world relevance. Quantitative methods --- statistics --- are needed. Examples:
\item Returns to education:
\begin{itemize}
\item What is the quantitative effect on student achievement of reducing class size by one student?
\item By how much does another year of education increase earnings?
\item How are these distributed among medical students, law students, economics students?
\item By how much does a Master's degree increase earnings above a Bachelor's degree?
\end{itemize}
\item Taxing tobacco:
\begin{itemize}
\item What is the price elasticity of cigarettes?
\item How does a one-dollar increase in cigarette taxes affect health outcomes?
\item By how much does a one-dollar increase in cigarette taxes reduce health expenditures?
\end{itemize}
\item Monetary policy:
\begin{itemize}
\item What is the effect on output growth of a one percentage point increase in the Fed's funds rate?
\end{itemize}
\item House prices:
\begin{itemize}
\item What is the effect on house prices of changes in zoning laws?
\end{itemize}
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Course Overview}
\begin{itemize}
\item This course is about using data to measure causal effects. For instance,
\begin{itemize}
\item Returns to education.
\item Cigarette prices.
\item Monetary policy.
\end{itemize}
\item Ideally, we would like an experiment.
\begin{itemize}  
\item what would be an experiment to estimate the effect of class size on standardized test scores?
\end{itemize}
\item But most of the data we have is not derived from designed experiments.
\item This course deals with challenges arising from using observational data to estimate causal effects:
\begin{itemize}  
\item Confounding effects (omitted factors).
\item Simultaneous causality.
\item Correlation does not imply causation!
\end{itemize}
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

