

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{What Is Econometrics?}
\begin{itemize}
\item Econometrics is a set of quantitative methods used for a variety of purposes. 
\begin{itemize}
\item To estimate relationships among economic variables.
\item To test economic theories.
\item To evaluate business practice. For instance, to identify the components of demand and costs, including labor costs.
\item To evaluate government policy, assess the effectiveness of government programs, subsidies, understand the sources of revenue.
\item To forecast future trends and fluctuations.
\item To estimate responses to future changes in policy and environmental circumstances.
\end{itemize}
\item Econometrics started as an applied branch of economics, but has since been applied to many other fields. In recent years, techniques developed in the context of machine learning are gradually being used in econometrics.
\item Econometrics is concerned with causal inference --- Machine learning may or may not be.
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Econometricians}
\begin{itemize}
\item Econometrics is specialized branch of statistics. Statistics originates in the seventeenth century. Econometrics is more recent, built on the foundations of modern statistics, in particular the work on inference of Ronald Fisher, Jerzy Neyman, Egon Pearson, and Abraham Wald.
\item Early pioneers include Jan Tinbergen, Ragnar Frisch, who won the first Nobel prize in economics in 1969. The other Nobel prize winners in the fields are: Simon Kuznets (1971), known for his applied work on national income, economic growth, and inequality; Tjalling Koopmans (1975); Trygve Haavelmo (1989); James Heckman, Daniel McFadden (2000); Robert Engle, Clive Granger (2003); Thomas Sargent, Christopher Sims (2011); Lars Peter Hansen (2013); Joshua Angrist, David Card, Guido Imbens (2021).
\item Other famous economists associated with the field of econometrics include (non-exhaustive list): Phillip and Sewell Wright; Halbert White; Peter Phillips; Bruce Hansen; Hashem Pesaran; Jeffrey Wooldridge; James Stock; Mark Watson; Whitney Newey; David Hendry; Jerry Hausman; James Poterba; Tim Bollerslev; George Tauchen; James Hamilton; Charles Manski; Pierre Perron; Francis Diebold; Zvi Griliches; Soren Johansen; Donald Rubin; Alan Krueger; Stephen Pischke; David Autor; Lawrence Katz; Richard Blundell; Stephen Bond; Manuel Arellano; John List; Christian Hansen; Victor Chernozhukov; Susan Athey \ldots\
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

