

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Problems and Applications}
\exercise{Stock \& Watson, Introduction (4th), Chapter~3, Empirical Exercise~1.}
The data file \textbf{CPS96\_15} contains data on full-time workers, ages 25-34, with a high school diploma or a B.A./B.S. as their highest degree. A detailed description is given in \textbf{CPS96\_15\_Description}, available on the website. Use these data to complete the following highlighted questions.
\color{blue}
\begin{enumerate}
\item 
  \begin{enumerate}
  \item Compute the sample mean for average hourly earnings ($\vn{AHE}$) in 1996 and 2015.
  \item Compute the sample standard deviation for $\vn{AHE}$ in 1996 and 2015.
  \item Construct a $95\%$ confidence interval for the population means of $\vn{AHE}$ in 1996 and 2015.
  \item Construct a $95\%$ confidence interval for the change in the population means of $\vn{AHE}$ between 1996 and 2015.
  \end{enumerate}
\item In 2015, the value of the Consumer Price Index (CPI) was $237.0$. In 1996, the value of the CPI was $156.9$. Repeat (a), but use $\vn{AHE}$ measured
in real 2015 dollars (\$2015); that is, adjust the 1996 data for the price inflation that occurred between 1996 and 2015.
\item If you were interested in the change in workers' purchasing power from 1996 to 2015, would you use the results from (a) or (b)? Explain.
\color{black} 
\item Using the data for 2015:
  \begin{enumerate}
  \item Construct a $95\%$ confidence interval for the mean of $\vn{AHE}$ for high
  school graduates.
  \item Construct a $95\%$ confidence interval for the mean of $\vn{AHE}$ for workers with a college degree.
  \item Construct a $95\%$ confidence interval for the difference between the two means.
  \end{enumerate} 
\item Repeat (d) using the 1996 data expressed in \$2015.
  \begin{enumerate}
  \item Using appropriate estimates, confidence intervals, and test statistics, answer the following questions:
  \item Did real (inflation-adjusted) wages of high school graduates increase from 1996 to 2015?
  \item Did real wages of college graduates increase?
  \item Did the gap between earnings of college and high school graduates increase? Explain.
  \end{enumerate}
\end{enumerate}
\end{enumerate}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

