---
title: "Empirical Exercise 2, Stock and Watson Chapter 8"
subtitle: "Econ 440 - Introduction to Econometrics"
author: "Patrick Toche, ptoche@fullerton.edu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup,  include=FALSE, eval=TRUE}
library(ggplot2)
library(broom)
library(dplyr)
library(tidyr)
options(digits=5)
```


## Empirical Exercise: Earnings
The data file **CPS2015** contains data for full-time, full-year workers, ages 25–34, with a high school diploma or B.A./B.S. as their highest degree. A detailed description is given in **CPS2015_Description**. In this exercise, you will investigate the relationship between a worker's age and earnings.


### Dataset:
```{r data, include=TRUE, eval=TRUE}
library(readxl)
df <- read_xlsx("CPS2015.xlsx", trim_ws=TRUE)
head(df)
```

## (a) 

Run a regression of average hourly earnings ($ahe$) on age ($age$), sex ($female$), and education ($bachelor$). If $age$ increases from 25 to 26, how are earnings expected to change? If $age$ increases from 33 to 34, how are earnings expected to change?

```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (b) 

Run a regression of the logarithm of average hourly earnings, $ln(ahe)$, on $age$, $female$, and $bachelor$. If $age$ increases from 25 to 26, how are earnings expected to change? If $age$ increases from 33 to 34, how are earnings expected to change?

```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (c) 

Run a regression of the logarithm of average hourly earnings, $ln(ahe)$, on $ln(age)$, $female$, and $bachelor$. If $age$ increases from 25 to 26, how are earnings expected to change? If $age$ increases from 33 to 34, how are earnings expected to change?

```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```


## (d)

Run a regression of the logarithm of average hourly earnings, $ln(ahe)$, on $age$, $age2$, $female$, and $bachelor$. If $age$ increases from 25 to 26, how are earnings expected to change? If $age$ increases from 33 to 34, how are earnings expected to change?

```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (e) 

Do you prefer the regression in (c) to the regression in (b)? Explain.


## (f)

Do you prefer the regression in (d) to the regression in (b)? Explain.

## (g) 

Do you prefer the regression in (d) to the regression in (c)? Explain.


## (h)

Plot the regression relation between $age$ and $ln(ahe)$ from (b), (c), and (d) for males with a high school diploma. Describe the similarities and differences between the estimated regression functions. Would your answer change if you plotted the regression function for females with college degrees?

```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```


## (i)

Run a regression of $ln(ahe)$ on $age$, $age2$, $female$, $bachelor$, and the interaction term $female * bachelor$. What does the coefficient on the interaction term measure? Alexis is a 30-year-old female with a bachelor's degree. What does the regression predict for her value of $ln(ahe)$? Jane is a 30-year-old female with a high school diploma. What does the regression predict for her value of $ln(ahe)$? What is the predicted difference between Alexis's and Jane's earnings? Bob is a 30-year-old male with a bachelor's degree. What does the regression predict for his value of $ln(ahe)$? Jim is a 30-year-old male with a high school diploma. What does the regression predict for his value of $ln(ahe)$? What is the predicted difference between Bob's and Jim's earnings?

```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (j)
Is the effect of Age on earnings different for men than for women? Specify and estimate a regression that you can use to answer this question.

```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (k)
Is the effect of Age on earnings different for high school graduates than for college graduates? Specify and estimate a regression that you can use to answer this question.

```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

## (l)
After running all these regressions (and any others that you want to run), summarize the effect of age on earnings for young workers.

```{r,  include=TRUE, eval=TRUE}
## fill this space with your code
```

