---
title: "Empirical Exercise 2, Stock and Watson Chapter 8"
subtitle: "Econ 440 - Introduction to Econometrics"
author: "Patrick Toche, ptoche@fullerton.edu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup,  include=FALSE, eval=TRUE}
library(ggplot2)
library(broom)
library(dplyr)
library(tidyr)
options(digits=5)
```

## Empirical Exercise: Earnings
The data file **CPS2015** contains data for full-time, full-year workers, ages 25–34, with a high school diploma or B.A./B.S. as their highest degree. A detailed description is given in **CPS2015_Description**. In this exercise, you will investigate the relationship between a worker's age and earnings.


### Dataset:
```{r data, include=TRUE, eval=TRUE}
library(readxl)
df <- read_xlsx("CPS2015.xlsx", trim_ws=TRUE)
head(df)
```

## (a)

Run a regression of average hourly earnings ($ahe$) on age ($age$), sex ($female$), and education ($bachelor$). If $age$ increases from 25 to 26, how are earnings expected to change? If $age$ increases from 33 to 34, how are earnings expected to change?

```{r,  include=TRUE, eval=TRUE}
m1 <- lm(ahe ~ age + female + bachelor, data=df)
summary(m1)
```

If $age$ increases by 1 year, $ahe$ is expected to increase by
$\$0.5313$, that is about $53$ cents per hour. Thus, the answer is the same for an increase from 25 to 26 and from 33 to 34. We can make a prediction for the values, rather than the increase. For convenience, we use R's built-in \texttt{predict} function.

```{r,  include=TRUE, eval=TRUE}
newdata <- data.frame(age=c(25,26), female=rep(0,2), bachelor=rep(0,2))
predict(m1, newdata)
```

As $age$ increases from 25 to 26, earnings increase from $15.327$ to $15.858$

```{r,  include=TRUE, eval=TRUE}
newdata <- data.frame(age=c(33,34), female=rep(0,2), bachelor=rep(0,2))
predict(m1, newdata)
```
As $age$ increases from 33 to 34, earnings increase from $19.577$ to $20.108$

As this is a repeat question, let's build a convenience function:

```{r,  include=TRUE, eval=TRUE}
lm_predict <- function(model, data){
  n <- length(data)
  newdata <- data.frame(age=data, 
                        female=rep(0,n), 
                        bachelor=rep(0,n), 
                        predicted=rep(1,n))
  newdata$ahe <- predict(model, newdata)
  newdata
}
lm_predict(m1, data=c(33,34))
```


## (b) 

Run a regression of the logarithm of average hourly earnings, $ln(ahe)$, on $age$, $female$, and $bachelor$. If $age$ increases from 25 to 26, how are earnings expected to change? If $age$ increases from 33 to 34, how are earnings expected to change?

```{r,  include=TRUE, eval=TRUE}
m2 <- lm(log(ahe) ~ age + female + bachelor, data=df)
summary(m2)
```
Because the regression involves non-linear transformations of the original variables, interpreting the coefficients requires great care. 

If $age$ increases by 1 year, $ln(ahe)$ is expected to increase by
$0.2419$ log-dollars, that is earnings $ahe$ is expected to increase by about $2.5\%$. This predicted effect is the same for an increase in $age$ from 25 to 26 and for an increase from 33 to 34. 

The predicted percentage change is computed as follows:
$$
\Delta\widehat{ln(ahe)} = 0.2419
\implies
\Delta\widehat{ahe} 
    \approx e^{0.2419}-1
    \approx 0.025
$$
A simple rule of thumb is $0.2419\to2.4\%$

This calculation may be done in R by extracting the coefficient from the linear model object. 
```{r,  include=TRUE, eval=TRUE}
exp(m2$coefficients[["age"]])-1
```
To preview the $m2$ object's structure, type \texttt{str(m2)} and then \texttt{str(m2\$coefficients)}.
```{r,  include=TRUE, eval=TRUE}
m2$coefficients[["age"]] * (log(34)-log(33))
```

Specifically, the predicted values of $ahe$ are:
```{r,  include=TRUE, eval=TRUE}
lm_predict(m2, data=c(25,26,33,34))
```




## (c) 

Run a regression of the logarithm of average hourly earnings, $ln(ahe)$, on $ln(age)$, $female$, and $bachelor$. If $age$ increases from 25 to 26, how are earnings expected to change? If $age$ increases from 33 to 34, how are earnings expected to change?

```{r,  include=TRUE, eval=TRUE}
m3 <- lm(log(ahe) ~ log(age) + female + bachelor, data=df)
summary(m3)
```
Both $ahe$ and $age$ appear in log form. As a result, it is no longer true that an increase in $age$ by 1 year increases $ahe$ uniformly: instead the effect on $ahe$ depends on $age$ and must be computed separately for each case.

If $age$ increases by 1 year, from 25 to 26, $ahe$ is expected to increase by about $2.8\%$:
$$
\Delta\widehat{ln(ahe)} = 0.7154 \cdot (\ln(26)-\ln(25))
    \approx 0.028
$$
This calculation follows from: 
```{r,  include=TRUE, eval=TRUE}
m3$coefficients[["log(age)"]] * (log(26)-log(25))
```
If $age$ increases by 1 year from 33 to 34, the increase in $ahe$ is smaller, at about $2.1\%$.
$$
\Delta\widehat{ln(ahe)} = 0.7154 \cdot (\ln(34)-\ln(33))
    \approx 0.021
$$

```{r,  include=TRUE, eval=TRUE}
m3$coefficients[["log(age)"]] * (log(34)-log(33))
```

Specifically, the predicted values of $ahe$ are:
```{r,  include=TRUE, eval=TRUE}
lm_predict(m3, data=c(25,26,33,34))
```



## (d)

Run a regression of the logarithm of average hourly earnings, $ln(ahe)$, on $age$, $age2$, $female$, and $bachelor$. If $age$ increases from 25 to 26, how are earnings expected to change? If $age$ increases from 33 to 34, how are earnings expected to change?

```{r,  include=TRUE, eval=TRUE}
m4 <- lm(log(ahe) ~ age + I(age^2) + female + bachelor, data=df)
summary(m4)
```
If $age$ increases by 1 year, $ahe$ is expected to increase by
$\$0.5313$, that is about $53$ cents per hour.

## (e) 

Do you prefer the regression in (c) to the regression in (b)? Explain.

The difference between the regressions in (b) and (c) is that $age$ is replaced by $log(age)$. The statistical analysis does not point to any major differences. The individual coefficients all have near-zero p-values. The regression's standard error is about the same. The regression's adjusted R-squared is about the same. The regression's p-value is near zero in both cases. The main reason for preferring regression (c) is that the coefficient on $log(age)$ can be interpreted as an elasticity. 

## (f)

Do you prefer the regression in (d) to the regression in (b)? Explain.

The difference between regression (b) and regression (d) is that the linear regressor $age$ is replaced by a quadratic polynomial in $age$ and $age^2$. The regression's standard error, adjusted R-squared, and p-value do not give arguments in favor of one or the other. However, since the individual coefficient on $age^2$ is significant at the $0.05$ level, suggesting the presence of some non-linearity, we prefer regression (d). 

## (g) 

Do you prefer the regression in (d) to the regression in (c)? Explain.

The difference between regression (c) and regression (d) is that $log(age)$ is replaced by a quadratic polynomial in $age$ and $age^2$. The regression's standard error, adjusted R-squared, and p-value do not give arguments in favor of one or the other. We prefer regression (c) because it is parsimonious and the coefficient may be interpreted as an elasticity.

The individual coefficient on $age^2$ has a smaller p-value than that on $log(age)$, but these coefficients cannot be compared since in regression (d) $age^2$ is picking up only the non-linear part of the causal relation between age and hourly earnings, while in regression (c) $log(age)$ is picking up all of the causal relation. See below for a graphical comparison of models (c) and (d).

## Remark
Differences in the R-squared across the models are non-existent or tiny and cannot be used to discriminate across models, despite what some model answers out there in the internets are saying.

## (h)

Plot the regression relation between $age$ and $ln(ahe)$ from (b), (c), and (d) for males with a high school diploma. Describe the similarities and differences between the estimated regression functions. Would your answer change if you plotted the regression function for females with college degrees?

## Create labels 
Create labels to replace "0" and "1" with "Male"/"Female" and "High School"/"Bachelor"
```{r,  include=TRUE, eval=TRUE}
female.labs <- c("0" = "Male", "1" = "Female")
bachelor.labs <- c("0" = "High School", "1" = "Bachelor")
```

## Augment the datasets 
Augment the datasets with the regression results (from broom package):
```{r,  include=TRUE, eval=TRUE}
dm1 <- augment(m1, data=df)
dm2 <- augment(m2, data=df)
dm3 <- augment(m3, data=df)
dm4 <- augment(m4, data=df)
```

## Plot of $log(ahe|female=0,bachelor=0)$ against $age$ 
### for regression (b)
```{r,  include=TRUE, eval=TRUE}
ggplot(data=dm2, aes(x=age, y=log(ahe))) + 
    geom_point() + 
    facet_grid(bachelor~female, 
               labeller=labeller(bachelor=bachelor.labs, female=female.labs)) +
    geom_smooth(method="lm", color="red",se=FALSE) +
    geom_segment(aes(xend=age, yend = .fitted), linetype="dashed") +
    ggtitle("log-linear model")
```


### for regression (c)
```{r,  include=TRUE, eval=TRUE}
ggplot(data=dm3, aes(x=age, y=log(ahe))) + 
    geom_point() + 
    facet_grid(bachelor~female, 
               labeller=labeller(bachelor=bachelor.labs, female=female.labs)) +
    geom_smooth(method="lm", color="red",se=FALSE) +
    geom_segment(aes(xend=age, yend = .fitted), linetype="dashed") +
    ggtitle("log-log model")
```

### for regression (d)
```{r,  include=TRUE, eval=TRUE}
ggplot(data=dm4, aes(x=age, y=log(ahe))) + 
    geom_point() + 
    facet_grid(bachelor~female, 
               labeller=labeller(bachelor=bachelor.labs, female=female.labs)) +
    geom_smooth(method="lm", color="red",se=FALSE) +
    geom_segment(aes(xend=age, yend = .fitted), linetype="dashed") +
    ggtitle("log-quadratic model")
```


## Discussion: Model (c) vs. Model (d)

### Plot of $log(ahe)$ against residuals
```{r,  include=TRUE, eval=TRUE}
ggplot(dm3, aes(log(age), .resid)) + 
  geom_point() + 
  geom_hline(yintercept=0, color="red", linetype='dashed') + 
  ggtitle("Residuals in the log-log model")
```

```{r,  include=TRUE, eval=TRUE}
ggplot(dm4, aes(age, .resid)) + 
  geom_point() + 
  geom_hline(yintercept=0, color="red", linetype='dashed') + 
  ggtitle("Residuals in the log-quadratic model")
```

### Plot of $log(ahe)$ against standardized residuals: qq plot
```{r,  include=TRUE, eval=TRUE}
ggplot(dm3, aes(sample=.std.resid)) + 
  geom_qq() +
  geom_qq_line(color="red") +
  ggtitle("qq-plot: standardized residuals in the log-log model")
```
```{r,  include=TRUE, eval=TRUE}
ggplot(dm4, aes(sample=.std.resid)) + 
  geom_qq() +
  geom_qq_line(color="red") +
  ggtitle("qq-plot: standardized residuals in the log-quadratic model")
```



## Models compared
```{r,  include=FALSE, eval=TRUE}
labs <- c("Female & High School", "Male & High School", "Female & Bachelor", "Male & Bachelor")
ggplot() + 
  geom_line(data=dm2, aes(x=age, y=.fitted, 
                          group=interaction(-female, bachelor), 
                          color=interaction(-female, bachelor)),
            size=1.5) +
  scale_color_manual(name="",labels=labs, values=c("red", "forestgreen", "orange", "blue")) +
  guides(color=guide_legend(reverse=TRUE)) +
  ggtitle("log-linear model") +
  theme_bw() -> p2
```

```{r,  include=FALSE, eval=TRUE}
ggplot() + 
  geom_line(data=dm3, aes(x=age, y=.fitted, 
                          group=interaction(-female, bachelor), 
                          color=interaction(-female, bachelor)),
            size=1.5) +
  scale_color_manual(name="",labels=labs, values=c("red", "forestgreen", "orange", "blue")) +
  guides(color=guide_legend(reverse=TRUE)) +
  ggtitle("log-log model") +
  theme_bw() -> p3
```

```{r,  include=FALSE, eval=TRUE}
ggplot() + 
  geom_line(data=dm4, aes(x=age, y=.fitted, 
                          group=interaction(-female, bachelor), 
                          color=interaction(-female, bachelor)),
            size=1.5) +
  scale_color_manual(name="",labels=labs, values=c("red", "forestgreen", "orange", "blue")) +
  guides(color=guide_legend(reverse=TRUE)) +
  ggtitle("log-quadratic model") +
  theme_bw() -> p4
```

```{r,  include=TRUE, eval=TRUE}
print(p2)
print(p3)
print(p4)
```


## (i)

Run a regression of $ln(ahe)$ on $age$, $age2$, $female$, $bachelor$, and the interaction term $female * bachelor$. 

```{r,  include=TRUE, eval=TRUE}
df$age2 <- (df$age)^2
df$femalebachelor <- df$female * df$bachelor
m5 <- lm(log(ahe) ~ age + age2 + female + bachelor + femalebachelor, data=df)
dm5 <- augment(m5, data=df)
summary(m5)
```

```{r,  include=FALSE, eval=TRUE}
library(stargazer)
stargazer(m2, m3, m4, m5, type="text", digits=2, out="reg_output.tex") 
```

### What does the coefficient on the interaction term measure? 

The coefficient shows the extra effect on earnings for females with a bachelor degree that is not captured by the average female and not captured by the average bachelor. 


### Alexis and Jane: 30-year old females
Alexis is a 30-year-old female with a bachelor's degree. What does the regression predict for her value of $ln(ahe)$? Jane is a 30-year-old female with a high school diploma. What does the regression predict for her value of $ln(ahe)$? What is the predicted difference between Alexis's and Jane's earnings? 
```{r,  include=TRUE, eval=TRUE}
alexis <- data.frame(age=30, age2=30^2, bachelor=1, female=1, femalebachelor=1)
log.ahe.alexis = predict(m5, newdata=alexis)
ahe.alexis = exp(log.ahe.alexis)
jane <- data.frame(age=30, age2=30^2, bachelor=0, female=1, femalebachelor=0)
log.ahe.jane = predict(m5, newdata=jane)
ahe.jane = exp(log.ahe.jane)
```

The predicted difference between Alexis's and Jane's earnings is about $8,053.7$ dollars:
```{r,  include=TRUE, eval=TRUE}
ahe.alexis - ahe.jane
```

### Bob and Jim: 30-year old males
Bob is a 30-year-old male with a bachelor's degree. What does the regression predict for his value of $ln(ahe)$? Jim is a 30-year-old male with a high school diploma. What does the regression predict for his value of $ln(ahe)$? What is the predicted difference between Bob's and Jim's earnings?
```{r,  include=TRUE, eval=TRUE}
bob <- data.frame(age=30, age2=30^2, bachelor=1, female=0, femalebachelor=1)
log.ahe.bob = predict(m5, newdata=bob)
ahe.bob = exp(log.ahe.bob)
jim <- data.frame(age=30, age2=30^2, bachelor=0, female=0, femalebachelor=0)
log.ahe.jim = predict(m5, newdata=jim)
ahe.jim = exp(log.ahe.jim)
```

The predicted difference between Bob's and Jim's earnings is about $9,742.1$ dollars:
```{r,  include=TRUE, eval=TRUE}
ahe.bob - ahe.jim
```

### Difference-in-Difference
The difference in the difference predicted effects is about $9,742.1-8,053.7=1,688.4$ dollars:
```{r,  include=TRUE, eval=TRUE}
(ahe.bob - ahe.jim) - (ahe.alexis - ahe.jane)
```


The predicted difference in log-earnings between Alexis and Jane is about $0.47559$ log-dollars. The predicted difference in log-earnings between Bob and Jim is about $0.47559$ log-dollars. That just happens to be the same! Thus, the difference in difference in log-earnings is about zero log-dollars. An earlier edition of Stock and Watson used a different sample dataset and reports a near-zero difference ($0.063$), which is consistent with our findings.

## (j)
Is the effect of Age on earnings different for men than for women? Specify and estimate a regression that you can use to answer this question.

A parsimonious model is the following:
```{r,  include=TRUE, eval=TRUE}
m6 <- lm(log(ahe) ~ log(age) + bachelor + log(age)*female + bachelor*female, data=df)
summary(m6)

library(car)
H0 <- c("female=0", "log(age):female", "bachelor:female=0")
tidy(linearHypothesis(m6, H0))
```
To test whether the effect of $age$ on $ahe$ is different for men and women, we test the hypothesis that all coefficients involving $female$ are jointly zero. This hypothesis can be firmly rejected since the joint p-value is essentially zero.


## (k)
Is the effect of Age on earnings different for high school graduates than for college graduates? Specify and estimate a regression that you can use to answer this question.

```{r,  include=TRUE, eval=TRUE}
m7 <- lm(log(ahe) ~ log(age) + female + log(age)*bachelor + female*bachelor, data=df)
summary(m7)

library(car)
H0 <- c("bachelor=0", "log(age):bachelor", "female:bachelor=0")
tidy(linearHypothesis(m7, H0))
```
To test whether the effect of $age$ on $ahe$ is different for university graduates and high-school graduates, we test the hypothesis that all coefficients involving $bachelor$ are jointly zero. This hypothesis can be firmly rejected since the joint p-value is essentially zero.

## (l)
After running all these regressions (and any others that you want to run), summarize the effect of age on earnings for young workers.

The following models fit about $21\%$ of the deviation of average hourly earnings from the sample mean. It's not immediately obvious that one should be preferred over the other.

### Log-quadratic model:
```{r,  include=TRUE, eval=TRUE}
m8 <- lm(log(ahe) ~ poly(age,2,raw=TRUE) + bachelor + female, data=df)
summary(m8)
dm8 <- augment(m8, data=df)
ggplot() + 
  geom_line(data=dm8, aes(x=age, y=.fitted, 
                          group=interaction(-female, bachelor), 
                          color=interaction(-female, bachelor)),
            size=1.5) +
  scale_x_continuous(breaks=seq(25,35,1)) +
  scale_color_manual(name="",labels=labs, values=c("red", "forestgreen", "orange", "blue")) +
  guides(color=guide_legend(reverse=TRUE)) +
  ggtitle("log-quadratic model") +
  theme_bw()
```

### Log-Log model:
```{r,  include=TRUE, eval=TRUE}
m9 <- lm(log(ahe) ~ log(age) + bachelor + female, data=df)
summary(m9)
dm9 <- augment(m9, data=df)
ggplot() + 
  geom_line(data=dm9, aes(x=age, y=.fitted, 
                          group=interaction(-female, bachelor), 
                          color=interaction(-female, bachelor)),
            size=1.5) +
  scale_x_continuous(breaks=seq(25,35,1)) +
  scale_color_manual(name="",labels=labs, values=c("red", "forestgreen", "orange", "blue")) +
  guides(color=guide_legend(reverse=TRUE)) +
  ggtitle("log-log model") +
  theme_bw()
```

