---
title: "Confidence Sets for Regression Coefficients"
subtitle: "Econ 440 - Introduction to Econometrics"
author: "Patrick Toche, ptoche@fullerton.edu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup,  include=FALSE, eval=TRUE}
library(ggplot2)
library(broom)
library(dplyr)
library(tidyr)
options(digits=5)
```

### Dataset:
```{r data, include=TRUE, eval=TRUE}
library(readxl)
df <- read_xlsx("caschool.xlsx", trim_ws=TRUE)
head(df)
```

## Regression
Rename variables to match textbook notation
```{r,  include=TRUE, eval=TRUE}
names(df)[names(df) == "testscr"] <- "TestScore"
names(df)[names(df) == "expn_stu"] <- "Expn"
names(df)[names(df) == "str"] <- "STR"
names(df)[names(df) == "el_pct"] <- "PctEL"
df$Expn <- df$Expn/1000

m1 <- lm(TestScore ~ STR + Expn + PctEL, data=df)
summary(m1)
```

## Confidence Ellipses
Draw confidence ellipses for regression coefficients on $STR$ and $Expn$.
```{r,  include=TRUE, eval=TRUE}
library(ellipse)
library(latex2exp)
conf.ellipse <- data.frame(ellipse(m1, which=c(2,3)))
ggplot(conf.ellipse, aes(x=STR, y=Expn)) + 
  geom_path() +
  geom_polygon(fill="cornflowerblue", alpha=0.5, color="darkblue", size=1) +
  scale_x_continuous(breaks=seq(-2,2,0.5), limits=c(-1.5,1.5)) + 
  scale_y_continuous(breaks=seq(-1,9,1)) + 
  geom_point(x=m1$coefficient[2], y=m1$coefficient[3], shape=3, size=2) +
  geom_point(x=0, y=0, shape=3, size=2) +
  labs(x="Coefficient on STR", y="Coefficient on Expn", 
       title="95% Confidence Set for Coefficients on STR and Expn") +
  annotate("segment", x=0.75, xend=0.05, y=0, yend=0) +
  annotate("segment", 
           x=m1$coefficient[2]+0.05, xend=0.75, 
           y=m1$coefficient[3]-0.1, yend=1) +
  annotate("text", x=0.8, y=1, 
           label=TeX("($\\hat{\\beta}_{1}, \\hat{\\beta}_{2})=(-0.29,3.87)$"), 
           parse=TRUE, hjust=0) +
  annotate("text", x=0.8, y=0, 
           label=TeX("($\\beta_{1}, \\beta_{2})=(0,0)$"), 
           parse=TRUE, hjust=0) +
  theme_classic()
ggsave(last_plot(), file="plot-regression-ellipses.pdf", width=8, height=5)
```

The $95\%$ confidence set Coefficient on $STR$ ($\beta_{1}$) and $Expn$ ($\beta_{2}$) is an ellipse. The ellipse contains the pairs of values of $\beta_{1}$ and $\beta_{2}$ that cannot be rejected using the F-statistic at the $5\%$ significance level. The point ($\beta_{1}, \beta_{2}$=(0,0)) is not contained in the confidence set, so the null hypothesis $H_{0}\colon \beta_{1}=0
\text{and} \beta_{2}=0$ is rejected at the $5\%$ significance level.



## Joint Hypothesis test
Joint test of hypothesis $H_{0}:$ $STR=0$ and $Expn=0$.
```{r,  include=TRUE, eval=TRUE}
library(car)
H0 <- c("STR=0", "Expn=0")
tidy(linearHypothesis(m1, H0))
```

