---
title: "Introduction to Linear Regression with R"
subtitle: "Econ 440 - Introduction to Econometrics"
author: "Patrick Toche, ptoche@fullerton.edu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
  html_document:
    df_print: paged
---

```{css my-header-colors, echo = FALSE}
.page-header {
    background-image: linear-gradient(120deg, #4863A0, #ADDFFF);
}
```

```{r setup,  include=FALSE, eval=TRUE}
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("tidyr")
# install.packages("ggthemes")
# # install.packages("prettydoc") # to theme knitr
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggthemes)
library(ggrepel)
options(digits=5)
```

### Rmarkdown themes

The overall theme of your notebook is controlled by the option `theme` in the `yaml` preamble. Supported themes include `cerulean`,  `cosmo`,  `flatly`,  `journal`,  `lumen`,  `paper`, `readable`, `sandstone`, `simplex`,  `spacelab`, `united`, and `yeti`. 

The highlighting theme is controlled by the option `highlight`, usually placed immediately below the theme. Supported styles include `default`, `tango`, `pygments`, `kate`, `monochrome`, `espresso`, `zenburn`, `haddock`, `breezedark`, and `textmate`.

See this [gallery](https://www.datadreaming.org/post/r-markdown-theme-gallery/) for examples. For more themes, you can use the extension package [prettydoc](https://cran.r-project.org/web/packages/prettydoc/vignettes/tactile.html). And you can also modify existing styles, or even create your own style from scratch, with `css` modifiers. See immediately below the `yaml` preamble of the source `Rmd` file for a simple example.


### Load dataset
```{r data, include=TRUE, eval=TRUE}
library(readxl)
df <- read_xlsx("Growth.xlsx", trim_ws=TRUE)
head(df)
```

### Make a scatterplot of average annual growth rate and average trade share:
```{r,  include=TRUE, eval=TRUE}
library(ggplot2)
df$country <- as.factor(df$country_name)
ggplot(data=df, aes(x=tradeshare, y=growth)) + geom_point()
```

### Detect the outlier: Print the country name
```{r,  include=TRUE, eval=TRUE}
ggplot(data=df, aes(x=tradeshare, y=growth, label=country)) + 
    geom_point() +
    geom_text()
```

### Detect the outlier: Print the country name + Tweak
```{r,  include=TRUE, eval=TRUE}
ggplot(data=df, aes(x=tradeshare, y=growth, label=country)) + 
    geom_point(col='blue', size=1) +
    geom_text(size=3)
```

### Detect the outlier: Use labels instead of plaint text
```{r,  include=TRUE, eval=TRUE}
ggplot(data=df, aes(x=tradeshare, y=growth, label=country)) + 
    geom_label(size=2)
```

### Detect the outlier: Highlight the variable name
```{r,  include=TRUE, eval=TRUE}
ggplot(data=df, aes(x=tradeshare, y=growth, label=country)) + 
    geom_text(size=2, aes(colour = I(ifelse(country == "Malta", "red", "black")))) 
```

### Detect the outlier: Avoid overlapping labels
```{r,  include=TRUE, eval=TRUE}
library(ggrepel)
ggplot(data=df, aes(x=tradeshare, y=growth, label=country)) + 
    geom_point(col="darkgreen", size=0.5) +
    geom_text_repel(aes(label=country), size=2) 
```

### Detect the outlier: Add a theme!
```{r,  include=TRUE, eval=TRUE}
ggplot(data=df, aes(x=tradeshare, y=growth, label=country)) + 
    geom_point(col="darkgreen", size=1) +
    geom_text_repel(aes(label=country), colour=I(ifelse(df$country == "Malta", "red", "black")), size=I(ifelse(df$country == "Malta", 4, 2))) +
    ggtitle("growth vs trade") +
    theme_wsj()
```

### Investigate correlation:
```{r,  include=TRUE, eval=TRUE}
cor.test(df$growth, df$tradeshare)
```

### Investigate linear regression:
```{r,  include=TRUE, eval=TRUE}
ols <- lm(growth ~ tradeshare, data=df)
summary(ols)
```

### Add regression line to the scatterplot:
```{r,  include=TRUE, eval=TRUE}
ggplot(data=df, aes(x=tradeshare, y=growth)) + 
    geom_point(col="darkgreen") +
    geom_smooth(method = "lm", se=FALSE)
```

### Add regression line to the scatterplot | confidence interval:
```{r,  include=TRUE, eval=TRUE}
ggplot(data=df, aes(x=tradeshare, y=growth)) + 
    geom_point(col="darkgreen") +
    geom_smooth(method = "lm", se=TRUE) +
    theme_bw()
```

### Regression without the outlier
```{r,  include=TRUE, eval=TRUE}
df2 <- subset(df, country != "Malta")
ggplot(data=df2, aes(x=tradeshare, y=growth)) + 
    geom_point(col="darkgreen") +
    geom_smooth(method = "lm", se=FALSE)
```

### Regression without the outlier
```{r,  include=TRUE, eval=TRUE}
ggplot(data=df, aes(x=tradeshare, y=growth)) + 
    geom_point(col="darkgreen") +
    geom_smooth(method = "lm", se=FALSE, data=df2) +
    labs(title = 'OLS is not robust to regression',
         caption="Regression line with outlier omitted", 
         x="trade share")
```

### Predict Malta
```{r,  include=TRUE, eval=TRUE}
# get regression coefficients
ols2 <- lm(growth ~ tradeshare, data=df2)
b0 <- coef(ols2)[1]
b1 <- coef(ols2)[2]
# get Malta trade share
x_obs <- df[df$country == "Malta", "tradeshare"]
y_hat <- b0 + b1 * x_obs
y_hat

# or use the built-in predict:
predict(ols2, newdata=x_obs)
```

### Prediction with/without Malta in sample, compared
```{r,  include=TRUE, eval=TRUE}
# observed value:
y_obs <- df[df$country == "Malta", "growth"]
y_obs
# predicted value with outlier:
predict(ols, newdata=x_obs)

# predicted value without outlier:
predict(ols2, newdata=x_obs)
```
