---
title: "Getting Started with R"
subtitle: "Econ 440 - Introduction to Econometrics"
author: "Patrick Toche, ptoche@fullerton.edu"
date: "`r format(Sys.time(), '%d %B %Y')`"
fontsize: 11pt
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: default
    fonttheme: professionalfonts
    slide_level: 2
    includes:
        in_header: ../beamer-header.tex
    keep_tex: yes
  prettydoc::html_pretty:
    theme: architect
    highlight: github
classoption: svgnames, handout, t
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, comment=NA, concordance=TRUE, cache=TRUE)
knitr::opts_chunk$set(dev=c('pdf','png'), fig.align='center', fig.height=c(4.5, 5), fig.width=c(8, 8), dpi=600, pdf.options(encoding="ISOLatin9.enc")) 
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
options(pillar.bold=TRUE, pillar.subtle_num=TRUE)
options(digits=3)
chunk_hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk=function(x, options) {
  x <- chunk_hook(x, options)
  paste0("\n \\", "footnotesize","\n\n", x, "\n\n \\normalsize")
})
library(knitr)
library(readxl)
library(scales)
library(dplyr)
library(tidyverse)
```



## Contents
1. Know your data types and data containers
2. Convert strings and factors to numeric values
3. Select/Create data inside a dataframe
4. Import/Export data
5. Quick sums
6. Split/Apply
7. Create functions



## Introduction
This document is work in progress. It will be updated as we proceed through the course, in response to questions that arise for particular tasks. It is designed to be permanently incomplete and permanently revised. Also, please use with care: Always back up your data!



## Know your data types
We are going to be dealing with the following data types: numbers, strings, booleans, factors, and containers filled with these, including: vectors/lists, named lists, matrices, dataframes, tibbles.

Start a new session, type the letter a into the console: you get an error!
```{r, include=TRUE, echo=TRUE, eval=FALSE, message=TRUE}
a
```

Now type the letter c into the console: you get something!
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
c
```

That's because in ``R`` the letter ``c`` is the name of a function ``c()``. That's unfortunate, but ``R`` was designed in the 1970s when very few people had a need of symbols beyond ``a`` and ``b``. Lol.



## Know your data types
Other than ``c()``, other names that are used by ``R``, and are therefore best avoided as names for other objects, include ``df`` and ``data``. But because they are names of functions, these names are actually still available for other objects. For instance,
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
# define c
c = 0

# check the value of c
c

# but c() is still available
c(0,1)
```
It's not too smart, but we do it all the time. Better people use ``C`` or ``DF``, because ``R`` is case sensitive and the upper-case symbols are available. 



## Know your data types
The following is definitely stupid:
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
# define c
c = function(x) NULL

# check c(0,1)
c()

# remove this before it ruins our life
rm(c)
```



## Know your data types
So let's investigate some data types:
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
# floating-point number aka num
a1 = 0
str(a1)

# string aka character chr
a2 = "0"
str(a2)

# boolean aka logical
a3 = TRUE
str(a3)
```
In ``R`` booleans are called ``logical`` vectors. 



## Know your data types
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
# integer aka int
a4 = 0L
str(a4)

# factor
a5 = factor(0, level=1)
str(a5)

# vector of factors
a6 = factor(c(0,1), levels=c(1,2))
str(a6)
```



## Know your data containers
We will be dealing mostly with 'vectors' and 'dataframes' In ``R`` lists are vectors. A vector is the most fundamental data container in ``R``. Even a single object is under the hood a vector with a single value. 
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
# a vector/list of numbers:
l1 = c(0, 1)
str(l1)

# a vector/list of integers:
l2 = c(0, 1L)
str(l2)

# a vector/list of booleans:
l3 = c(FALSE, TRUE)
str(l3)
```



## Know your data containers
In ``R`` named lists have a purpose similar to 'dictionaries' in ``Python``.
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
# a named list:
l4 = c("name1" = 1, "name2" = TRUE)
str(l4)
```

Every object stored in a named list has a name and a value:
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
# get the name and value with []
l4["name1"]

# get the value with [[]]
l4[["name1"]]
```



## Know your data containers
The easiest way to create a dataframe is from several vectors.
```{r, include=TRUE, echo=TRUE, eval=TRUE}
# define band member, the date they joined/quit the band, and their net worth
# data from a quick web search - not meant to be accurate!
member <- c('Brian', 'Mick', 'Keith', 'Stewart', 'Bill', 'Charlie', 'Ronnie')
start <- as.Date(c('1962-07-12', '1962-07-12','1962-07-12','1962-07-12',
                   '1962-12-07','1963-02-02','1976-04-23'))
today <- as.character(Sys.Date())
end <- as.Date(c('1969-06-09', today, today, 
                 '1963-05-01', '1992-12-01', '2021-08-24', today))
worth <- c(10, 500, 500, 1, 80, 250, 200)
df <- rolling.stones <- data.frame(member, start, end, worth)
str(rolling.stones)

# add information about the units:
rolling.stones$currency <- "$"  # R is clever that way
```



## Know your data containers
We can output the core structure of a dataframe with the ``dput()`` function.
```{r, include=TRUE, echo=TRUE, eval=TRUE}
dput(df)
```
This is useful to share the dataframe without creating a ``csv`` file or similar. It is the best way to share toy data on websites like [stackoverflow](https://stackoverflow.com/).

You can convert the dataframe to a ``tibble`` using the ``as_tibble()`` function:
```{r, include=TRUE, echo=TRUE, eval=TRUE}
dput(as_tibble(df))
```
which shows that the ``tibble`` does not contain more information than the ``dataframe``.



## Convert strings and factors to numeric values
Conversion across types must be done with care, obviously. 

Functions such as ``as.numeric()``, ``as.character()`` are 'vectorized', which means they can be applied to a vector/list and also to a single column in a dataframe.
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
# convert a number ot a string:
as.character(0)

# convert a string to a number:
as.numeric("0")

# convert a string to an integer:
as.integer("0")
as.integer("0.5")

# convert a vector of booleans:
as.numeric(c(TRUE, FALSE))
```



## Convert strings and factors to numeric values
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
# convert a factor to number:
f = factor(c(0,1), levels=c(1,2))
as.character(f)
as.numeric(as.character(f))

# convert a column of a dataframe and reassign back to dataframe
df$worth <- as.integer(df$worth)
str(df)
```



## Select data inside a dataframe
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
# get a column
df$member

# get one row
df[2,]

# get several rows
df[2:3,]

# get a row by criterion
df[df$worth >= 300,]
```



## Add data to a dataframe
We can create categorical variables as follows:
```{r, include=TRUE, echo=TRUE, eval=TRUE}
df$poor <- df$worth < 100
df$middle <- df$worth >= 100 & df$worth < 300
df$rich <- df$worth >= 300
```

We can create a duration variable as follows:
```{r, include=TRUE, echo=TRUE, eval=TRUE}
df$duration <- round(100*as.integer((df$end-df$start))/as.integer(max(df$end-df$start)))
```
This calculates the interval between the start and end dates, calculates the maximum time interval in the datasets, and for each interval calculates the percentage it represents. 
```{r, include=TRUE, echo=TRUE, eval=TRUE}
df
```



## Use data in a dataframe
The correlation between net worth and the duration variable is positive, as expected:
```{r, include=TRUE, echo=TRUE, eval=TRUE}
cor(df$duration, df$worth)
```

And significant:
```{r, include=TRUE, echo=TRUE, eval=TRUE}
cor.test(df$duration, df$worth)
```



## Import data
The most common data format is ``csv`` --- comma separated values. 

The most commonly needed options are ``stringsAsFactors=FALSE`` to prevent ``R`` from converting strings to factors and the option to use or discard header information, with ``header=FALSE`` or ``header=TRUE``:
```{r, include=TRUE, echo=TRUE, eval=FALSE}
# basic import:
d <- read.csv("Age_HourlyEarnings.csv", stringsAsFactors=FALSE)

# import with more control:
d <- read.csv("Age_HourlyEarnings.csv", stringsAsFactors=FALSE, 
              skip=1, na.strings="NA", header=TRUE, strip.white=TRUE)
```

There are many ways to import excel data. One option is to use the ``read_xlsx()`` function from the ``readxl`` package.
```{r, include=TRUE, echo=TRUE, eval=FALSE}
library(readxl)
d <- read_xlsx("Age_HourlyEarnings.xlsx", col_names=TRUE, skip=1, trim_ws=TRUE)
```



## Export data
To save a dataframe as a ``csv`` file:
```{r, include=TRUE, echo=TRUE, eval=FALSE}
write.csv(df, file = "filename.csv", row.names=FALSE)
```
where you typically do not want to save the row names.

To save a dataframe as a ``xlsx`` file, among many options, you can use the ``write_xlsx()`` function from the ``writexl`` package:
```{r, include=TRUE, echo=TRUE, eval=FALSE}
library(writexl)
write_xlsx(df, file = "filename.csv", row.names=FALSE)
```

You can also save the data in the native ``RData`` format. This format will preserve information about data types, information that will otherwise be lost with the ``csv`` and ``xlsx`` formats. The only downside of the ``RData`` format is that it cannot be imported by spreadsheets and may not be easily read by other software, whereas the ``csv`` format is universal and the ``xlsx`` format nearly universal. 
```{r, include=TRUE, echo=TRUE, eval=FALSE}
save(df, file = "filename.RData")
```



## Export data
It is possible to import data from other formats, including ``Stata``, ``SPSS``, ``SAS``. One option, among several, comes from the ``Hmisc`` and ``foreign`` packages. Examples:
```{r, include=TRUE, echo=TRUE, eval=FALSE}
library(foreign)
df.stata <- read.dta("stata-format.dta")

library(Hmisc)
df.spss <- spss.get("spss-format.por", use.value.labels=TRUE)

library(Hmisc)
df.sas <- sasxport.get("sas-format.xpt")
```



## Quick sums
``R`` knows how to sum the elements of a vector:
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
x = c(1,2,3,4,5)
sum(x)

x = c(TRUE, TRUE)
sum(x)
```

Other operations are possible in the same way:
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
x = c(TRUE, FALSE)
mean(x)
```

To sum across the columns or rows of a dataframe or matrix, the ``colSums()`` and ``rowSums()`` are optimized for speed. Similar are ``colMeans()`` and ``rowMeans()``.
```{r, include=TRUE, echo=TRUE, eval=TRUE}
m = matrix(1:9, nrow = 3)
colSums(m)
rowMeans(m)
```



## Split/Apply
To apply a particular calculation to a subset of the data, the ``split()`` function is convenient.  
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
# split the data: poor vs the rest
split(df, df$poor)

# apply a calculation: compute the mean for each group
sapply(split(df$worth, df$poor), mean)
```



## Functions
It is often convenient to create convenience functions. Ah yes. 

A population standard deviation does not exist in base ``R``. For quick check, ``?sd``: The Help page for ``sd`` states "Like `var` this uses denominator n-1."

Let's roll our own:
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
std <- function(x) sqrt(sum((x-mean(x))^2)/length(x))
std(df$worth)
```
Note that we divide by `length(x)` and not ``length(x)-1``

Functions that stretch over several lines use an opening and closing curly brace:
```{r, include=TRUE, echo=TRUE, eval=TRUE, message=TRUE}
weighted.std <- function(x,w) {
    sqrt(sum(w*(x-weighted.mean(x,w))^2)/sum(w))
}
std(df$worth)
```


## Dataframe V. List
There is a subtle difference between the two lines of code below. 
```{r,  include=TRUE, eval=TRUE}
library(tibble)
d <- as_tibble(df)
d[d$poor, "member"]
d[d$poor,]$member
```

They yield the same result here. Execute code snippets to see how `R` works:
So what is the difference? Try `str()` on the resulting object to see.
```{r,  include=TRUE, eval=TRUE}
# the tibble
str(d[d$poor, "member"])
str(d[d$poor,]$member)
# the dataframe
str(df[df$poor, "member"])
str(df[df$poor,]$member)
```

Beware of tibbles, sometimes. See the difference?
```{r,  include=TRUE, eval=TRUE}
# sum
sum(d[d$poor, "worth"])
sum(d[d$poor,]$worth)
# mean
mean(d[d$poor, "worth"])
mean(d[d$poor,]$worth)
```




## END OF DOCUMENT
This document ends here. Please check back later for updates.


